Integrating random choice method with SPH : a new scheme with adaptive artificial viscosity and shock capturing ability

### 2018-01-08 JCP-Manuscript.tex

\modulolinenumbers[5]
\journal{Journal of Computational Physics}
%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%
%% Numbered
\bibliographystyle{model1-num-names}
%% Numbered without titles
%\bibliographystyle{model1a-num-names}
%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}
%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}
%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}
%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\begin{frontmatter}
\title{Integrating random choice method with SPH : a new scheme with adaptive artificial viscosity and shock capturing ability}
%\tnotetext[mytitlenote]{Can add  title note here}
%% Group authors per affiliation:
\author{Zhixuan Cao}
\author{Abani Patra \fnref{mycorrespondingauthor}}
\address{Department of Mechanical and Aerospace Engineering,The State University of New York at Buffalo, Amherst, NY, 14260-4200, USA}
\author{E. Bruce Pitman}
\address{Department of Materials Design and Innovation, The State University of New York at Buffalo, Amherst, NY, 14260-4200, USA}
%\fntext[myfootnote]{Since 1880.}
%% or include affiliations in footnotes:
%\author[mymainaddress,mysecondaryaddress]%{Elsevier Inc}
%\ead[url]{www.elsevier.com}
%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Abani Patra}
\ead{abani@buffalo.edu}
%\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
%\address[mysecondaryaddress]{360 Park Avenue South, New York}
\begin{abstract}
Standard smoothed particle hydrodynamics (SPH) employs an artificial viscosity, defined explicitly by constant coefficients, to properly capture hydrodynamic shock waves.
Such methodology usually provides dissipation more than need and smears the discontinuity if the artificial viscosity coefficients are not tuned.
Excessively large artificial viscosity might also damp generation of turbulence by fluid instability.
We present in this article a novel SPH scheme with adaptive artificial viscosity, in which the artificial viscosity is determined implicitly based on solution of a local Riemann problem. Different from Godunov SPH (GSPH) method, which utilizing solution of local Riemann problem at an imaginary interface, the new method randomly sample the solution of Riemann problem mimicking the random choice method (RCM). One dimensional shock tube tests demonstrate several attractive features of this new scheme. First of all, it introduces less but sufficient artificial dissipation for shock capturing compared with classical SPH and GSPH. The pressure ``wiggle" around contact discontinuity, which persists in SPH results, is ameliorated by the new scheme. It also shows higher order of convergence than GSPH.
%when the same construction of local Riemann problem and Riemann solvers are adopted by both methods. 
Combining of RCM with SPH overcomes the inability of classical RCM method in solving multi-dimensional unsteady non-linear systems. Three dimensional time dependent compressible Euler equations are solved for simulating a free jet flow, demonstrating that the desirable less dissipation property of the new scheme persists in multi-dimensional applications.
\end{abstract}
\begin{keyword}
Smoothed Particle Hydrodynamics (SPH), Random Choice Method (RCM), Riemann Solver, Hyperbolic PDEs, Artificial viscosity
\end{keyword}
\end{frontmatter}

\linenumbers
\section{Introduction}
Smoothed particle hydrodynamics (SPH) \citep{gingold1977smoothed,lucy1977numerical} is a mesh-free particle method based on Lagrangian formulation, and has been widely applied to different areas in engineering and science. SPH has both advantages and disadvantages with respect to other numerical techniques. One main advantage is that it automatically adapts to following dynamic flows and arbitrary geometries. Another advantage is that it needs less programming effort to include more physics. In addition, SPH particles only interact with neighboring particles, which avoids massive global communication in parallel implementation of SPH.
Historically, SPH has difficulty in capturing discontinuities, such as shocks or contact discontinuities. Inviscid description of fluids becomes invalid at shock fronts, where specific entropy of the fluids increases through the conversion of mechanical energy into internal energy. Most SPH codes have included an artificial viscosity term explicitly, both for stabilizing the simulation and for handling shock discontinuities by dissipating local velocity differences and converting them into heat \citep{monaghan1983shock, monaghan1997sph, klapp2012strong}. Without knowing the minimum amount of dissipation for shock capturing, such methodology usually provides dissipation more than need and smears the discontinuity if the artificial viscosity coefficients are not tuned. The method of using constant artificial viscosity might also introduce spurious effects at region away from shocks, for instance, causing an unphysical damping of turbulent mixing (as shown, for example, by \citet{borgani2012hydrodynamic}) or spurious
shearing torques in rotating flows \citep{flebbe1994smoothed}. Selective application of artificial viscosity techniques, such as the time dependent viscosity \citep{morris1997switch, dolag2005turbulent} and shock-indicator-based viscosity \citep{cullen2010inviscid} have been tried for counteracting this. 
\citet{sigalotti2008adaptive} improved the standard SPH near sharp discontinuities adopting an adaptive density kernel estimation (ADKE) techniques. However, the parameters for ADKE have to be tuned to the problem at hand. Hence it is hard to implement ADKE in real applications.

@@

The random choice method (RCM) was introduced by \citet{glimm1965solutions} for the construction of solutions of systems of nonlinear hyperbolic conservation laws. \citet{chorin1976random} developed the RCM as a practical computational method for solutions to one dimensional (1D) Euler equations of gas dynamics. Chorin's work was followed by numerous researchers, with applications and further improvements \citep{sod1977numerical,concus1979numerical,colella1982glimm, freistuhler1992numerical}. In essence, the solution is advanced in time by a sequence of operations that includes the solution of Riemann problems and a stochastic sampling procedure. When applied to one dimensional time dependent Euler equations \citep{colella1982glimm}, diffusive errors from spatial averaging are completely eliminated and hence discontinuities, such as shocks or material interface, can be resolved as true discontinuities.
For unsteady flows in genuine two (or higher) space dimensions, however, such very desirable property of RCM do not seem to persist \citep{colella1982glimm}. Implementing of RCM within the framework of SPH only involves sampling of solutions of Riemann problem in 1D local coordinate system no matter what is the dimensionality of the problem. Hence desirable property of RCM method in 1D should be retained in higher dimensional implementations within the framework of SPH. This hypothesis is confirmed by a three-dimensional (3D) free jet flow simulation.
%This need to be proven ---> But at least, if the good property of preserving true discontinuity is still valid in 1D shock tube problem, it should be preserved in multiple-dimensional implementations. Because in current framework of RSPH, no matter what is the dimension of the problem, the techniques is only implemented in a one-dimensional scenario.

In this investigation, we introduce a new method, which will be referred to as RSPH in later paragraph, by integrating RCM method within framework of SPH. 
The new scheme inherits the attraction of classical RCM introducing less but sufficient artificial viscosity and causing less smearing of discontinuities than other methods for 1D problems. 
For multi-dimensional problem, RCM method is implemented within one-dimensional local coordinate system, established for pair of interacting particles. So no multi-dimensional coordinate is involved even for multi-dimensional problems. Thereby, the attractive properties of less dissipation showing in one-dimensional case is preserved in multi-dimensional implementations. In this way, the RSPH overcomes RCM's fundamental limitation of losing attractive property showing in 1D in higher dimensional application.

RSPH has the following advantages over standard SPH:
\begin{itemize}
\item RSPH is able to capture discontinuities with less smearing.
\item The RSPH is able to adaptively adjust equivalent artificial viscosity coefficients, assigning larger artificial viscosity coefficients around the shock and small artificial viscosity coefficients otherwhere.
\item RSPH is able to alleviate pressure ``wiggle" at the contact discontinuity, a common issue in classical SPH.
\end{itemize}
Compared with GSPH, which is also based on solution of local Riemann problem, RSPH introduces less but sufficient dissipation than GSPH, especially in the region away from shocks. Such feature of RSPH is especially beneficial for simulating of scenarios involving turbulence mixing, for which excessive dissipation would suppress mixing. In addition, RSPH also shows faster convergence rate than GSPH.

@@

Our investigation focuses on implementation of the RSPH schemeit in solving inviscid compressile Euler equations .  
\begin{align}
\dfrac{\partial \rho}{\partial t} + \nabla \cdot \left(\rho \textbf{v} \right) = 0 \label{eq:gov-cs-rho} \\
\dfrac{\partial \rho \textbf{v}}{\partial t} + \nabla \cdot \left(\rho \textbf{v} \textbf{v} + p\textbf{I}\right) = 0 \label{eq:gov-cs-v} \\
\dfrac{\partial \rho E}{\partial t} + \nabla \cdot \left[\left(\rho E + p \right)\textbf{v}\right] = 0 \label{eq:gov-cs-e}
\end{align}
where $\rho$ is the density, $\textbf{v}$ is the velocity, $\textbf{I}$ is a unit tensor.
$E = e + K $ is the total energy which is a summation of kinetic energy $K$ and internal energy $e$.
The system of equations is closed by equation of state (EOS) for idea gas.
\begin{equation}
p = \left(\gamma - 1\right)\rho e \label{eq:EOS}
\end{equation}
with $\gamma=1.4$.

For any SPH scheme, governing equations in Lagrange form are needed. By deducting kinetic energy from energy equation, subtracting mass conservation from momentum equation, combining transient term and advection term into material derivative (for any function $A$, material derivative is defined as $\frac{D A}{Dt} = \frac{\partial A}{\partial t} + \textbf{v} \cdot \nabla A$) term, the governing equations are put into the final form. 
\begin{align}
\dfrac{D \rho}{D t} + \rho \nabla \cdot \textbf{v} = 0 \label{eq:gov-nc-rho}\\
\dfrac{D \textbf{v}}{D t} + \dfrac{\nabla P}{\rho} =0 \label{eq:gov-nc-v}\\
\dfrac{D e}{D t} + \dfrac{P \nabla \cdot \textbf{v}}{\rho} = 0 \label{eq:gov-nc-e}
\end{align}
%\cutet{puri2014comparison} propose an addition of an artificial heating term to the GSPH scheme to eliminate unphysical spikes in the thermal energy at the contact discontinuity.  --> It is worthwhile to check whether RSPH is able to avoid this phenomena without using his techniques.
%Another issue of RCM: the very desirable properties of RCM: no diffusive errors from spatial averaging do not seem to persist for flows in two space dimensions. For two dimensional steady supersonic flow, one of the space variables can be treated as a time-like one and RCM can be applied in a way similar to that for one-dimensional unsteady flow.
%--Then what about RSPH?

@@

where, $a$ is the SPH particle index, $\textbf{v}_{a b} = \textbf{v}_a - \textbf{v}_b$. $\Pi$ is an artificial viscosity term.

@@

$\alpha = 1$ and $\beta = 2$ are recommended \citep{monaghan2005smoothed} for best results and mostly adopted. In the numerical test section, these two parameters are varied to show the effect of them. $\eta$ is usually taken as 0.1 to prevent singularities.

In Eq. (\ref{eq:ns-sph-d}) - Eq. (\ref{eq:ns-sph-e}), $w_{a b}\left(h\right)$ is weighting function and is a concise form of $w\left(\textbf{x}_a - \textbf{x}_b, h\right)$. From here on, we will use this concise form.
There is a wide variety of possible weighting functions, such as spline functions (with different orders) and Gaussian functions. We adopt a truncated Gaussian function as the weighting function for standard SPH, GSPH and RSPH.
\begin{equation}
w\left(\textbf{x} - \textbf{x} \prime \right) = 
\begin{cases} 
      \dfrac{1}{\left(h \sqrt{\pi}\right)^d} exp \left[- \left(\dfrac{\textbf{x} - \textbf{x} \prime}{h} \right)^2 \right] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
      0 & \text{Otherwise}
\end{cases}
\label{eq:SPH-kernel}
\end{equation}
where $d$ is number of dimensions.
The derivative of the weighting function is:
\begin{equation}
\nabla w\left(\textbf{x} - \textbf{x} \prime \right) = 
\begin{cases} 
      -2\left(\dfrac{\textbf{x} - \textbf{x} \prime}{h}\right) \dfrac{1}{\left(h \sqrt{\pi}\right)^d} exp \left[- \left(\dfrac{\textbf{x} - \textbf{x} \prime}{h}\right)^2 \right] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
      0 & \text{Otherwise}
\end{cases}
\label{eq:SPH-kernel-gradient}
\end{equation}

@@

\subsection{GSPH method}
Here we present essential formulations for solving Euler equations with GSPH. The density field is updated in the same way as updating of density in SPH, according to Eq. (\ref{eq:ns-sph-d}). Updating of velocity and internal energy in GSPH is different from these in SPH. 
Derivation of discretized momentum and energy equation in first GSPH paper \citep{inutsuka2002reformulation} is based on the following discrete consistency identities: 
\begin{equation}
1=\sum_{b} \frac{m_{b}}{\rho(\textbf{x})}w(\textbf{x} - \textbf{x}_{b}, h)
\label{eq:GSPH-basic1}
\end{equation}
\begin{equation}
0=\sum_{b} m_{b} \nabla \frac{w(\textbf{x} - \textbf{x}_{b}, h)}{\rho(\textbf{x})}
\label{eq:GSPH-basic2}
\end{equation}
Lengthy algebraic manipulations \citep{inutsuka2002reformulation,iwasaki2011smoothed} lead to the discretized momentum and energy equations:
\begin{equation}
\ddot{\textbf{x}}_{a} = <\dfrac{d \textbf{v}_{a}}{dt}>= -\sum_{b} m_{b} p_{a b}^{\ast} \left[\frac{1}{\rho_{a}^2} \nabla w_{a b}(h_{a}) + \frac{1}{\rho_{b}^2} \nabla w_{a b}(h_{b}) \right]
\label{eq:gov-gsph-v-simple-form}
\end{equation}
\begin{equation}
<\dfrac{d e_{a}}{dt}>= - \sum_{b} m_{b} p_{a b}^{\ast} [\textbf{v}_{a b}^{\ast} - \dot{\textbf{x}}_{a}^{\ast}] \left[\frac{1}{\rho_{a}^2} \nabla w_{a b}(h_{a}) + \frac{1}{\rho_{b}^2} \nabla w_{a b}(h_{b}) \right]
\label{eq:gov-gsph-e-simple-form}
\end{equation}
where, $p_{a b}^{\ast}$ and $\textbf{v}_{a b}^{\ast}$ are obtained by solving a 1D Riemann problem constructed based on physical quantity of particle $a$ and particle $b$. The starred quantity  are interpolated value of the solution of the Riemann problem at an imaginary interface $q^{\ast}_{ab}$.
$\dot{\textbf{x}}_{a}^{\ast}$ is time centered velocity of particle $a$:
\begin{equation}
\dot{\textbf{x}}_{a}^{\ast} = <\textbf{v}_{a}> + \frac{\Delta t}{2} \ddot{\textbf{x}}_{a}
\end{equation}

The difference between this formulation and classical SPH formulations come from two sources: 
\begin{itemize}
\item the averaged (in certain sense) pressure and the effect of artificial viscosity is replaced by  $p^{\ast}$, obtained based on a local Riemann solver.
\item A time centered velocity is used, which satisfies the following identities approximately. 
\begin{equation}
\left( \textbf{v}_{a b}^{\ast} - \dot{\textbf{x}}_{a}^{\ast} \right) \approx -0.5 \textbf{v}_{a b}
\label{eq:SPH-GSPH-difference2}
\end{equation}
Where $\textbf{v}_{a b}$ is defined as $\textbf{v}_a - \textbf{v}_b$.
\end{itemize}
Actually, Godunov's idea of using Riemann solver can be combined with other SPH discretization formulations \citep{cha2003implementations}.

To conduct a well controlled comparison, the constructing of Riemann problem, Riemann solvers and projection of picked solution of Riemann problem back to global coordinate system are exactly the same in GSPH and RSPH, details of which will be described in detail in the next section. The way how solutions of Riemann problems are used is the only difference between GSPH and RSPH in our test. While RSPH takes solution of Riemann problem based on random sampling, solution of Riemann problem at an imaginary interface is picked up in GSPH, either based on interpolation of specific volume \citep{inutsuka2002reformulation} or simply assuming the middle point as the interface \citep{cha2003implementations}.

\section{RSPH method}
RSPH shares many commonality with SPH and GSPH, especially with GSPH. We adopt the same density updating formulation (see Eq. (\ref{eq:ns-sph-d})) for density updating in SPH, GSPH and RSPH. The same kernel function is used for all schemes. The equation for updating particle position (see Eq. (\ref{eq:SPH-update-pos})) is also the same for all schemes. Discretized governing equation for RSPH and GSPH are exactly the same while the way to determine $p^{\ast}$ and $\textbf{v}^{\ast}$ are different. We focus more on where RSPH is different from SPH and GSPH in this section.

>>>>

\modulolinenumbers[5]
\journal{Journal of Computational Physics}
%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%
%% Numbered
\bibliographystyle{model1-num-names}
%% Numbered without titles
%\bibliographystyle{model1a-num-names}
%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}
%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}
%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}
%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}
%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}
%% `Elsevier LaTeX' style
%\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\begin{frontmatter}
\title{Integrating random choice method with SPH : a new scheme with adaptive artificial viscosity and shock capturing ability}
%\tnotetext[mytitlenote]{Can add  title note here}
%% Group authors per affiliation:
\author{Zhixuan Cao}
\author{Abani Patra \fnref{mycorrespondingauthor}}
\address{Department of Mechanical and Aerospace Engineering,The State University of New York at Buffalo, Amherst, NY, 14260-4200, USA}
\author{E. Bruce Pitman}
\address{Department of Materials Design and Innovation, The State University of New York at Buffalo, Amherst, NY, 14260-4200, USA}
%\fntext[myfootnote]{Since 1880.}
%% or include affiliations in footnotes:
%\author[mymainaddress,mysecondaryaddress]%{Elsevier Inc}
%\ead[url]{www.elsevier.com}
%\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
\fntext[mycorrespondingauthor]{abani@buffalo.edu}
%\cortext[mycorrespondingauthor]{abani@buffalo.edu}
%\ead{abani@buffalo.edu}
%\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
%\address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
Standard smoothed particle hydrodynamics (SPH) employs an artificial viscosity, defined explicitly by constant coefficients, to properly capture hydrodynamic shock waves.
Such methodology usually provides dissipation more than need and smears the discontinuities if the artificial viscosity coefficients are not tuned.
Excessively large artificial viscosity might also damp generation of turbulence by fluid instability.
We present in this article a novel SPH scheme with adaptive artificial viscosity, in which the artificial viscosity is determined implicitly based on solution of a local Riemann problem. Different from Godunov SPH (GSPH) method, which utilizing solution of local Riemann problem at an imaginary interface, the new method randomly sample the solution of Riemann problem mimicking the random choice method (RCM). One dimensional shock tube tests demonstrate several attractive features of this new scheme. First of all, it introduces less but sufficient artificial dissipation for shock capturing compared with classical SPH and GSPH. The pressure ``wiggle" around contact discontinuity, which persists in SPH results, is ameliorated by the new scheme. The new method also shows higher order of convergence than GSPH.
%when the same construction of local Riemann problem and Riemann solvers are adopted by both methods. 
Classical RCM method would lose its desirable property of low numerical dissipation when solving multi-dimensional unsteady non-linear systems. Combining of RCM with SPH overcomes the shortcoming  of classical RCM method. Three dimensional time dependent compressible Euler equations are solved by the new scheme for simulating a free jet flow, demonstrating that the desirable less dissipation property of the new scheme persists in multi-dimensional applications.
\end{abstract}
\begin{keyword}
Smoothed Particle Hydrodynamics (SPH), Random Choice Method (RCM), Riemann Solver, Artificial viscosity, Hyperbolic PDEs
\end{keyword}
\end{frontmatter}

\linenumbers
\section{Introduction}
Smoothed particle hydrodynamics (SPH) \citep{gingold1977smoothed,lucy1977numerical} is a mesh-free particle method based on Lagrangian formulation and has been widely applied to different areas in engineering and science. SPH has both advantages and disadvantages with respect to other numerical techniques. One main advantage is that it automatically adapts to following dynamic flows and arbitrary geometries. Another advantage is that it needs less programming effort to include more physics. In addition, SPH particles only interact with neighboring particles, which avoids massive global communication in parallel implementation of SPH.
Historically, SPH has difficulty in capturing discontinuities, such as shocks or contact discontinuities. Inviscid description of fluids becomes invalid at shock fronts, where specific entropy of the fluids increases through the conversion of mechanical energy into internal energy. Most SPH codes have included an artificial viscosity term explicitly, both for stabilizing the simulation and for handling shock discontinuities by dissipating local velocity differences and converting them into heat \citep{monaghan1983shock, monaghan1997sph, klapp2012strong}. Without knowing the minimum amount of dissipation for shock capturing, such methodology usually provides dissipation more than need and smears the discontinuity if the artificial viscosity coefficients are not tuned. The method of using constant artificial viscosity might also introduce spurious effects at region away from shocks, for instance, causing an unphysical damping of turbulent mixing (as shown, for example, by \citet{borgani2012hydrodynamic}) or spurious
shearing torques in rotating flows \citep{flebbe1994smoothed}. Selective application of artificial viscosity techniques, such as the time dependent viscosity \citep{morris1997switch, dolag2005turbulent} and shock-indicator-based viscosity \citep{cullen2010inviscid} have been tried for counteracting this. 
\citet{sigalotti2008adaptive} improved the standard SPH near sharp discontinuities adopting adaptive density kernel estimation (ADKE) techniques. However, the parameters for ADKE have to be tuned to the problem at hand \citep{puri2014comparison}. Hence it is hard to implement ADKE in real applications.

@@

The random choice method (RCM) was introduced by \citet{glimm1965solutions} for the construction of solutions of systems of nonlinear hyperbolic conservation laws. \citet{chorin1976random} developed the RCM as a practical computational method for solutions to one dimensional (1D) Euler equations of gas dynamics. Chorin's work was followed by numerous researchers, with applications and further improvements \citep{sod1977numerical,concus1979numerical,colella1982glimm, freistuhler1992numerical}. In essence, the solution is advanced in time by a sequence of operations that include solving of Riemann problems and a stochastic sampling of the solution. See \citep{toro2013riemann} for details. When applied to one dimensional time dependent Euler equations \citep{colella1982glimm}, diffusive errors from spatial averaging are completely eliminated and hence discontinuities, such as shocks or material interface, can be resolved as true discontinuities.
For unsteady flows in genuine two (or higher) space dimensions, however, such very desirable property of RCM do not seem to persist \citep{colella1982glimm}.
%This need to be proven ---> But at least, if the good property of preserving true discontinuity is still valid in 1D shock tube problem, it should be preserved in multiple-dimensional implementations. Because in current framework of RSPH, no matter what is the dimension of the problem, the techniques is only implemented in a one-dimensional scenario.

In this investigation, we introduce a new method, which will be referred to as RSPH in later paragraph, by integrating RCM method within framework of SPH. 
The new scheme inherits the attraction of classical RCM, introducing less but sufficient artificial viscosity and causing less smearing of discontinuities than other methods for 1D problems. Implementing of RCM within the framework of SPH only involves sampling of solutions of Riemann problem in 1D local coordinate system no matter what is the dimensionality of the problem. That is to say, no multi-dimensional coordinate is involved even for multi-dimensional problems. Thereby, the desirable property of RCM method in 1D should be retained in higher dimensional implementations within the framework of SPH. This hypothesis is confirmed by a three-dimensional (3D) free jet flow simulation.

To summarize, RSPH has the following advantages over standard SPH:
\begin{itemize}
\item RSPH is able to capture discontinuities with less smearing.
\item The RSPH is able to adaptively adjust equivalent artificial viscosity coefficients, assigning larger artificial viscosity coefficients around the shock and small artificial viscosity coefficients otherwhere.
\item RSPH is able to alleviate pressure ``wiggle" at the contact discontinuity, a common issue in classical SPH.
\end{itemize}
Compared with GSPH, which is also based on solution of local Riemann problems, RSPH introduces less but sufficient dissipation than GSPH, especially in the region away from shocks. Such feature of RSPH is especially beneficial for simulating of scenarios involving turbulence mixing, for which excessive dissipation would suppress mixing. In addition, RSPH also shows faster convergence rate than GSPH.

@@

Our investigation implements RSPH scheme in solving inviscid compressile Euler equations as the application case.  
\begin{align}
\dfrac{\partial \rho}{\partial t} + \nabla \cdot \left(\rho \textbf{v} \right) = 0 \label{eq:gov-cs-rho} \\
\dfrac{\partial \rho \textbf{v}}{\partial t} + \nabla \cdot \left(\rho \textbf{v} \textbf{v} + p\textbf{I}\right) = 0 \label{eq:gov-cs-v} \\
\dfrac{\partial \rho E}{\partial t} + \nabla \cdot \left[\left(\rho E + p \right)\textbf{v}\right] = 0 \label{eq:gov-cs-e}
\end{align}
where $\rho$ is the density, $\textbf{v}$ is the velocity, $\textbf{I}$ is a unit tensor.
$E = e + K $ is the total energy which is a summation of kinetic energy $K$ and internal energy $e$.
The system of equations is closed by equation of state (EOS) for idea gas.
\begin{equation}
p = \left(\gamma - 1\right)\rho e \label{eq:EOS}
\end{equation}
with $\gamma=1.4$.

For any SPH scheme, governing equations in Lagrange form are needed. By deducting kinetic energy from energy equation, subtracting mass conservation from momentum equation, combining transient term and advection term into material derivative (for any function $A$, material derivative is defined as $\frac{D A}{Dt} = \frac{\partial A}{\partial t} + \textbf{v} \cdot \nabla A$) term, the governing equations are converted into the Lagrange form. 
\begin{align}
\dfrac{D \rho}{D t} + \rho \nabla \cdot \textbf{v} = 0 \label{eq:gov-nc-rho}\\
\dfrac{D \textbf{v}}{D t} + \dfrac{\nabla P}{\rho} =0 \label{eq:gov-nc-v}\\
\dfrac{D e}{D t} + \dfrac{P \nabla \cdot \textbf{v}}{\rho} = 0 \label{eq:gov-nc-e}
\end{align}
%\cutet{puri2014comparison} propose an addition of an artificial heating term to the GSPH scheme to eliminate unphysical spikes in the thermal energy at the contact discontinuity.  --> It is worthwhile to check whether RSPH is able to avoid this phenomena without using his techniques.
%Another issue of RCM: the very desirable properties of RCM: no diffusive errors from spatial averaging do not seem to persist for flows in two space dimensions. For two dimensional steady supersonic flow, one of the space variables can be treated as a time-like one and RCM can be applied in a way similar to that for one-dimensional unsteady flow.
%--Then what about RSPH?

@@

where, $a$ is the SPH particle index, $\textbf{v}_{a b} = \textbf{v}_a - \textbf{v}_b$. The smoothing length $h$ is taken as $h=\frac{h_a + h_b}{2}$. $\Pi$ is an artificial viscosity term.

@@

$\alpha = 1$ and $\beta = 2$ are recommended \citep{monaghan2005smoothed} for best results and mostly adopted. In the numerical test section, these two parameters are varied to show the effect different amount of numerical dissipation on simulation results. $\eta$ is usually taken as 0.1 to prevent singularities.

In Eq. (\ref{eq:ns-sph-d}) - Eq. (\ref{eq:ns-sph-e}), $w_{a b}\left(h\right)$ is weighting function and is a concise form of $w\left(\textbf{x}_a - \textbf{x}_b, h\right)$. We will use this concise form throughout this paper.
There is a wide variety of possible weighting functions, such as spline functions (with different orders) and Gaussian functions. We adopt a truncated Gaussian function as the weighting function for standard SPH, GSPH and RSPH.
\begin{equation}
w\left(\textbf{x} - \textbf{x} \prime, h \right) = 
\begin{cases} 
      \dfrac{1}{\left(h \sqrt{\pi}\right)^d} exp \left[- \left(\dfrac{\textbf{x} - \textbf{x} \prime}{h} \right)^2 \right] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
      0 & \text{Otherwise}
\end{cases}
\label{eq:SPH-kernel}
\end{equation}
where $d$ is number of dimensions.
The derivative of the weighting function is:
\begin{equation}
\nabla w\left(\textbf{x} - \textbf{x} \prime, h \right) = 
\begin{cases} 
      -2\left(\dfrac{\textbf{x} - \textbf{x} \prime}{h}\right) \dfrac{1}{\left(h \sqrt{\pi}\right)^d} exp \left[- \left(\dfrac{\textbf{x} - \textbf{x} \prime}{h}\right)^2 \right] &  \vert \textbf{x} - \textbf{x} \prime \vert \leq 3h\\
      0 & \text{Otherwise}
\end{cases}
\label{eq:SPH-kernel-gradient}
\end{equation}

@@

\subsection{GSPH method} \label{sec:GSPH-method}
Here we present essential formulations for solving Euler equations with GSPH. The density field is updated in the same way as updating of density in SPH, according to Eq. (\ref{eq:ns-sph-d}). Updating of velocity and internal energy in GSPH is different from these in SPH. 
Derivation of discretized momentum and energy equation in first GSPH paper \citep{inutsuka2002reformulation} is based on the following discrete consistency identities: 
\begin{equation}
1=\sum_{b} \frac{m_{b}}{\rho(\textbf{x})}w(\textbf{x} - \textbf{x}_{b}, h)
\label{eq:GSPH-basic1}
\end{equation}
\begin{equation}
0=\sum_{b} m_{b} \nabla \frac{w(\textbf{x} - \textbf{x}_{b}, h)}{\rho(\textbf{x})}
\label{eq:GSPH-basic2}
\end{equation}
Lengthy algebraic manipulations \citep{inutsuka2002reformulation,iwasaki2011smoothed} lead to the discretized momentum and energy equations:
\begin{equation}
\ddot{\textbf{x}}_{a} = <\dfrac{d \textbf{v}_{a}}{dt}>= -\sum_{b} m_{b} p_{a b}^{\ast} \left[\frac{1}{\rho_{a}^2} \nabla w_{a b}(h_{a}) + \frac{1}{\rho_{b}^2} \nabla w_{a b}(h_{b}) \right]
\label{eq:gov-gsph-v-simple-form}
\end{equation}
\begin{equation}
<\dfrac{d e_{a}}{dt}>= - \sum_{b} m_{b} p_{a b}^{\ast} [\textbf{v}_{a b}^{\ast} - \dot{\textbf{x}}_{a}^{\ast}] \left[\frac{1}{\rho_{a}^2} \nabla w_{a b}(h_{a}) + \frac{1}{\rho_{b}^2} \nabla w_{a b}(h_{b}) \right]
\label{eq:gov-gsph-e-simple-form}
\end{equation}
where, $p_{a b}^{\ast}$ and $\textbf{v}_{a b}^{\ast}$ are obtained by solving a 1D Riemann problem constructed based on physical quantity of particle $a$ and particle $b$. The starred quantity are interpolated value of the solution to the Riemann problem at an imaginary interface. More discussion about the imaginary interface is in section \ref{sec:Picking-up-single-state}.
$\dot{\textbf{x}}_{a}^{\ast}$ is time centered velocity of particle $a$:
\begin{equation}
\dot{\textbf{x}}_{a}^{\ast} = <\textbf{v}_{a}> + \frac{\Delta t}{2} \ddot{\textbf{x}}_{a}
\end{equation}

The difference between this formulation and classical SPH formulations come from two sources: 
\begin{itemize}
\item the averaged (in certain sense) pressure and the effect of artificial viscosity is replaced by  $p^{\ast}$, obtained based on a local Riemann solver.
\item a time centered velocity is used, which satisfies the following identities approximately. 
\begin{equation}
\left( \textbf{v}_{a b}^{\ast} - \dot{\textbf{x}}_{a}^{\ast} \right) \approx -0.5 \textbf{v}_{a b}
\label{eq:SPH-GSPH-difference2}
\end{equation}
%Where $\textbf{v}_{a b}$ is defined as $\textbf{v}_a - \textbf{v}_b$.
\end{itemize}
Actually, Godunov's idea of using Riemann solver can be combined with other SPH discretization formulations as well \citep{cha2003implementations}.

More technical details of GSPH are in section \ref{sec:RSPH-method}. The construction of Riemann problem, Riemann solvers and projection of picked solution of the Riemann problem back to global coordinate system are exactly the same in GSPH and RSPH. They are described in the next section. The way how to pick up the starred value from solution of the Riemann problem is the only difference between GSPH and RSPH. Detailed discussion is presented in section \ref{sec:Picking-up-single-state}.

\section{RSPH method} \label{sec:RSPH-method}
RSPH shares many commonalities with SPH and GSPH, especially with GSPH. We adopt the same density updating formulation (see Eq. (\ref{eq:ns-sph-d})) for density updating in SPH, GSPH and RSPH. The same kernel function is used for all schemes. The equation for updating particle position (see Eq. (\ref{eq:SPH-update-pos})) is also the same for all schemes. Discretized governing equations for RSPH and GSPH are exactly the same as well. The common palces shared with classical SPH and GSPH are unnecessary to repeat in this section. We focus more on where RSPH is different from SPH and GSPH.

Both GSPH and RSPH borrow ideas of shock capturing techniques from mesh based method. There is a great deal of commonality between Godunov's method and RCM for mesh based method. Both of them utilize the solution of a local Riemann problem. The difference is that the RCM pick up a single state contained in the local solutions ``randomly" while Godunov's method take the integral average. 
Integrating of Godunov's idea within SPH framework (the GSPH method), however, does not take the integral average at all. Instead, a single state of the local Riemann solution at an imaginary interface is picked up and used to upgrade physical quantities through discretized governing equations (see Eq. (\ref{eq:gov-gsph-v-simple-form}) - Eq. (\ref{eq:gov-gsph-e-simple-form})). The RSPH method, proposed in this paper mimicking classical RCM, picks up the single state by stochastic sampling. So the difference between RSPH and GSPH is how to pick up the single state from solution of the Riemann problem.
